---
title: "Image Compression"
author: "Ngoc Duong"
date: "5/30/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jpeg)
library(ISLR)
library(factoextra)
library(gridExtra)
library(corrplot)
library(RColorBrewer)
library(gplots)
```


### Hierarchical Clustering

```{r}
#load data
data("USArrests")

data = USArrests
```

a) Cluster states using complete linkage and Euclidean distance

```{r}
hc.complete <-hclust(dist(data), method = "complete")

fviz_dend(hc.complete, k = 4,
          cex = 0.3,
          palette = "jco",
          color_labels_by_k = TRUE,
          rect = TRUE, rect_fill = TRUE, 
          rect_border = "jco",labels_track_height = 2.5)
```


b) Cut the dendrogram at height that results in 3 distinct clusters

```{r}
fviz_dend(hc.complete, k = 3,
          cex = 0.3,
          palette = "jco",
          color_labels_by_k = TRUE,
          rect = TRUE, rect_fill = TRUE, 
          rect_border = "jco",labels_track_height = 2.5)
```

Look at the states that are in each clusters

```{r}
ind3.complete <-cutree(hc.complete, 3)

#cluster 1
data[ind3.complete==1,]

#cluster 2
data[ind3.complete==2,]

#cluster 3
data[ind3.complete==3,]
```


c) Hierarchically cluster the states using complete linkage and Euclidean distance after scaling the variables to have SD = 1

```{r}
dat1 <-scale(data)
```

```{r}
hc.complete.scale <-hclust(dist(dat1), method = "complete")

fviz_dend(hc.complete.scale, k = 3,
          cex = 0.3,
          palette = "jco",
          color_labels_by_k = TRUE,
          rect = TRUE, rect_fill = TRUE, 
          rect_border = "jco",labels_track_height = 2.5)
```

```{r}
ind3.complete.scale <-cutree(hc.complete.scale, 3)

#cluster 1
dat1[ind3.complete.scale==1,]

#cluster 2
dat1[ind3.complete.scale==2,]

#cluster 3
dat1[ind3.complete.scale==3,]
```

d) We observe that the cluster memberships changed after scaling the variables. The variables should be scaled before the inter-observation dissimilarities are computed in order to ensure equal weights given to every variable in $\mathbf{X}$ despite their different scales.

### PCA

```{r}
img <- readJPEG('piggy.jpg')

dim(img)

r <- img[,,1]
g <- img[,,2]
b <- img[,,3]

img.r.pca <- prcomp(r, center = FALSE)
img.g.pca <- prcomp(g, center = FALSE)
img.b.pca <- prcomp(b, center = FALSE)

rgb.pca <- list(img.r.pca, img.g.pca, img.b.pca)

# Approximate X with XV_kV_k^T
compress <- function(pr, k)
{
  compressed.img <- pr$x[,1:k] %*% t(pr$rotation[,1:k])
  compressed.img
}

# Using first 20 PCs
pca20 <- sapply(rgb.pca, compress, k = 20, simplify = "array")

writeJPEG(pca20, "pca20.jpeg")

# Try to increase the number of PCs!
pca50 <- sapply(rgb.pca, compress, k = 50, simplify = "array")

writeJPEG(pca50, "pca50.jpeg")

pca100 <- sapply(rgb.pca, compress, k = 100, simplify = "array")

writeJPEG(pca100, "pca100.jpeg")

pca200 <- sapply(rgb.pca, compress, k = 200, simplify = "array")

writeJPEG(pca200, "pca200.jpeg")
```

